Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-18 bbbq_article.tex
Range in percentage of TMH in human proteome is caused by the use of multiple TMH prediction tools. Fix #180

Transmembrane helices (TMHs) are hydrophobic 
as they need to span the hydrophobic lipid bilayer of cellular membranes, 
and they consist of an alpha helix of usually 23 amino acids. 
TMHs are common structures in the proteins of humans and pathogens. 
For example, 
15-39% 
\richel{Why wide range? Issue 180}
of the proteins in humans 
are predicted to be a membrane protein 
containing at least one TMH (\cite{ahram2006estimation}). 
TMHs can also be predicted from a protein sequence 
with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).

@@

In the final set of experiments, we addressed the question whether TMHs would be more evolutionary conserved than soluble proteins regions by comparing the occurrences of SNPs in the genome coding for TMHs and soluble protein regions. 
We obtained 1,129 gene names associated with the phrase 'membrane protein'.
These genes are linked to 5,203 proteins
\richel{GvdB: Why so many? Isoforms?}, of which 2,754
are predicted to be transmembrane proteins.
We obtained 4,811 SNPs that resulted in an
amino acids substitution, of which 2,568 were located 
in predicted transmembrane proteins.
Per protein, we predicted the percentages of the gene coding for TMHs and soluble regions
and calculated the percentage of amino acids substitutions
occurring in TMHs.
Per protein, this percentage was plotted as a function of the percentage TMH in figure
\ref{fig:f_snps_found_and_expected}.
We performed a linear regression analysis on the data, with and without
membrane-associated proteins \richel{Dit is nog niet goed uitgelegd}, and calculated the 95\%
confidence interval.
Since the regression curve clearly deviated from the line of equality,
which would be expected if SNPs occur just as likely within transmembrane and soluble protein domains,
this suggests that variations occur more often in soluble 
than in transmembrane protein regions.

@@

  Now it is just a list of potential reasons why the results might be flawed. 
  While this is in principle OK, 
  it is also nitpicking as some things affect only a limited number of 
  proteins (like selenomethionines) 
  or you’d expect changes in the other direction (wrong prediction of TMHs for bacteria).

  So offer interpretations. 
  The 2nd part of the introduction would fit here well. I can also help with this.
}

>>>>

Transmembrane helices (TMHs) are hydrophobic 
as they need to span the hydrophobic lipid bilayer of cellular membranes, 
and they consist of an alpha helix of usually 23 amino acids. 
TMHs are common structures in the proteins of humans and pathogens. 
Different TMH prediction tools estimate
that 15-39% of the proteins in humans 
contain at least one TMH (\cite{ahram2006estimation}). 
TMHs can also be predicted from a protein sequence 
with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).

@@

In the final set of experiments, we addressed the question whether TMHs would be more evolutionary conserved than soluble proteins regions by comparing the occurrences of SNPs in the genome coding for TMHs and soluble protein regions. 
We obtained 1,129 gene names associated with the phrase 'membrane protein'.
These genes are linked to 5,203 proteins
\richel{GvdB: Why so many? Isoforms?}, of which 2,754
are predicted to be transmembrane proteins.
We obtained 4,811 SNPs that resulted in an
amino acids substitution, of which 2,568 were located 
in predicted transmembrane proteins.
Per protein, we predicted the percentages of the gene coding for TMHs and soluble regions
and calculated the percentage of amino acids substitutions
occurring in TMHs.
Per protein, this percentage was plotted as a function of the percentage TMH in figure
\ref{fig:f_snps_found_and_expected}.
We performed a linear regression analysis on the data, with and without
membrane-associated proteins \richel{GvdB: explain clearer}, and calculated the 95\%
confidence interval.
Since the regression curve clearly deviated from the line of equality,
which would be expected if SNPs occur just as likely within transmembrane and soluble protein domains,
this suggests that variations occur more often in soluble 
than in transmembrane protein regions.

@@

  Now it is just a list of potential reasons why the results might be flawed. 
  While this is in principle OK, 
  it is also nitpicking as some things affect only a limited number of 
  proteins (like selenoproteins) 
  or you’d expect changes in the other direction (wrong prediction of TMHs for bacteria).

  So offer interpretations. 
  The second part of the introduction would fit here well. I can also help with this.
}
